\section{Problem Formulation}

In the considered OIRS-assisted VLC system, the objective is to determine the
optimal orientation of the optical intelligent reflecting surface (OIRS) that
maximizes the achievable data rate (ADR). The OIRS orientation is characterized
by the rotation angles $\gamma$ and $\omega$, which directly determine the
effective optical channel gain $z(\gamma,\omega)$ and, consequently, the
communication performance.

Based on the capacity lower bound for intensity-modulation/direct-detection
(IM/DD) channels, the achievable data rate is expressed as
\begin{equation}
\mathrm{ADR}(\gamma,\omega)
= \frac{B}{2}\log_{2}\!\left[
1+\frac{e}{2\pi}
\left(\frac{P\,R_{\mathrm{PD}}\,z(\gamma,\omega)}{Q}\right)^{2}
\right],
\label{eq:adr}
\end{equation}
where $B$ denotes the modulation bandwidth, $P$ is the transmitted optical
power, $R_{\mathrm{PD}}$ is the photodetector responsivity, $Q$ denotes the
noise standard deviation at the receiver, and $z(\gamma,\omega)$ is the
effective optical channel gain corresponding to the selected OIRS orientation.

Accordingly, the OIRS orientation optimization problem is formulated as
\begin{equation}
\begin{aligned}
(\mathrm{P1}):\qquad
\max_{\gamma,\,\omega}\quad
& \mathrm{ADR}(\gamma,\omega) \\
\text{s.t.}\qquad
& \gamma_{\min}\leq\gamma\leq\gamma_{\max}, \\
& \omega_{\min}\leq\omega\leq\omega_{\max}.
\end{aligned}
\label{eq:p1}
\end{equation}
The optimal orientation angles are denoted by $(\gamma^{*},\omega^{*})$.

Since the optimization variables $\gamma$ and $\omega$ are continuous and the
relationship between the OIRS orientation and the effective channel gain is
highly nonlinear and non-convex, obtaining a closed-form solution to Problem
(P1) is computationally intractable. Therefore, the optimization problem is
solved using deep reinforcement learning (DRL), in which the optimal OIRS
orientation is learned through direct interaction with the communication
environment.

At each decision step, the DRL agent observes the current system state,
consisting of the transmitter, OIRS, and receiver geometry together with the
current OIRS orientation, and outputs the continuous action vector
\begin{equation}
\mathbf{a}_t=
\begin{bmatrix}
\gamma_t \\
\omega_t
\end{bmatrix},
\end{equation}
where $\gamma_t$ and $\omega_t$ denote the orientation angles selected at the
$t$-th decision step. Based on the selected action, the environment evaluates
the effective channel gain $z(\gamma_t,\omega_t)$ and computes the achievable
data rate using \eqref{eq:adr}. The resulting ADR is returned as the reward
signal, enabling the DRL agent to iteratively learn the optimal orientation
angles that maximize the communication performance.

To solve Problem (P1), the proposed Enhanced Continuous Deep Policy
Optimization (ECDPO) algorithm is compared with four benchmark DRL algorithms,
namely Deep Deterministic Policy Gradient (DDPG), Soft Actor--Critic (SAC),
Proximal Policy Optimization (PPO), and Continuous Action Policy Gradient
(CAPG). All algorithms are evaluated under the same system model, state
representation, action space, and reward function to ensure a fair comparison
in terms of convergence speed, learning stability, and achieved ADR. The
comparative analysis identifies the most effective DRL algorithm for optimal
OIRS orientation in VLC systems.
